---
title: "simpleRegress"
author: "Fernando D Velasquez"
date: "April 12, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
bears$SEX <-as.factor(bears$SEX)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
WeightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(WeightModel)
summary.lm(WeightModel)
```

```{r}
ChestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(ChestModel)
summary.lm(ChestModel)
```
  
```{r}
AgeModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(AgeModel)
summary.lm(AgeModel)
```
```{r}
HeadLengthModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~NECK,data=bears)
abline(HeadLengthModel)
summary.lm(HeadLengthModel)
```

```{r}
HeadWthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(HeadWthModel)
summary.lm(HeadWthModel)
```


##SUMMARY
  
We can conclude after creating 5 different models that the best model for predicting Bears length is Head length. The models has an r squared of 0.8434 making it have an 84% of error been explained by the mean model.The residual error is 4.235 for head length.
